Using Genetic Algorithms to Optimize the Analogue Method for Precipitation Prediction in the Swiss Alps

### 2017-01-18 Optimization case study - JoH/elsarticle-template.tex
Changes to check.

Parallel analyses showed that the analogy of circulation is incomplete, and that geopotential heights still contain relevant information that can improve the statistical relationship. Therefore, a third, followed by a fourth predictor were added (still only geopotential heights). There was no constraint on the predictors, so that the same pressure level could be selected more than once. Further improvements were found on the performance score, both for the CP and the VP, confirming that this additional information was beneficial for the quality of the prediction.

@@

The number of predictors (still only geopotential heights) was then successively increased up to ten, considering the weighting of the analogy criteria values. The addition of predictors globally improved the prediction skill (for both the CP and the VP) only up to four predictors (Figure \ref{fig:figure_nb_levels}). Afterwards, the score on the VP was more variable, showing eventually even a decrease, which revealed an over-parametrization of the method, and thus a lack of robustness. After four predictors, the score on the CP did not increase substantially, and could even present some local decrease due to increasing difficulty for the optimizer to converge. Selecting four predictors (geopotential heights) was considered optimal for this case study, since the gain in CRPSS was significant, and the model remained relatively simple. It cannot be excluded that another number would prevail in another region than the upper Rh\^{o}ne catchment, under other meteorological conditions, or with another reanalysis dataset.

@@

The chosen method for the atmospheric circulation analogy, based on four predictors of geopotential heights, and that will be named 4Zo (o for optimized), was made of the following degrees of freedom:

@@

Globally, the same pattern could be observed for the VP (Figure \ref{fig:crossing_4Zo} bottom), but in this case, minor improvements could occur when crossing the parameters, due to the presence of other events in the VP that might have been better predicted by a different parameter set. The relatively small differences in score between parameterizations indicated that even though the parameters might differ significantly, the performance might not be drastically affected. Even a change in the pressure level did not mean a radical drop of the score value. A different parametrization may lead to a distinct selection of analogue days, and thus to an improvement of the prediction under certain weather conditions at the expense of others.

@@

Two methods were assessed: one with two moisture predictors (moisture index on two pressure levels or at two different hours) named 4Zo-2MIo, and one with four moisture predictors named 4Zo-4MIo. When introducing two predictors for the moisture analogy, the number of degrees of freedom raised to 42, and to 54 with four predictors.

@@

\subsection{Results for 4Zo-2MIo and 4Zo-4MIo methods}

Due to significant similarities between the results from 4Zo-2MIo and 4Zo-4MIo, the latter will only be partly shown.

@@

The CRPSS scores of the optimized 4Zo-2MIo method are provided in Figure \ref{fig:figure_crpss_4Zo-2HIo} and amounted in average to 40~\% (CP) and 40.3~\% (VP), compared to 35.2~\% (CP) and 36.2~\% (VP) for the reference method on the moisture analogy optimized with the sequential procedure. The score for the 4Zo-4MIo method were relatively close, with 40.6~\% for both the CP and the VP (not shown as very similar repartition as Figure \ref{fig:figure_crpss_4Zo-2HIo}).

As for 4Zo, the 4Zo-2MIo and 4Zo-4MIo methods presented larger improvements in the prediction for significant rainfall (thresholds P\(\geq\)1~mm, P\(\geq\)0.1\(\cdot\)P10 and P\(\geq\)0.5\(\cdot\)P10). The improvement were relatively similar for 4Zo-2MIo (Table \ref{table:scores_thresholds_4Zo-2MIo}) and 4Zo-4MIo (not shown), with slightly superior scores for the latter on small precipitation (P\(\geq\)1~mm : 17.7~\% and 13.0~\%) and extremes (P\(\geq\)0.5\(\cdot\)P10 : 23.7~\% and 29.2~\%).

The parameters cross-compatibility has also been assessed for the methods with moisture variables, and are shown in Figure \ref{fig:crossing_4Zo-2MIo} for the method 4Zo-2MIo (not shown for 4Zo-4MIo, but similar).

@@

When optimizing a method made of two levels of analogy, the introduction of moisture variables in the second level had an influence on the parameter values of the first level. It means that the two levels of analogy bring complementary information, and are thus not independent. This is first visible on the number $N_{1}$ of analogues to be selected on the first level, and on the selection of the pressure levels for the circulation analogy.

As for the sequential procedure, the optimal value of $N_{1}$ increased when adding a second level of analogy (Figure \ref{fig:figure_nb_analogs}). One can also notice that the optimal number of analogues $N_{2}$ for the second level of analogy of 4Zo-2MIo was slightly inferior to $N_{1}$ from 4Zo, but very close. There was globally a tendency between the optimal analogue number values of both methods: $N_{1}$ of the 4Zo method, and $N_{1}$ and $N_{2}$ of 4Zo-2MIo tended to be higher or lower altogether for a given region.

The optimal final numbers of analogues did not vary much: $23 \leq N_{1} \leq 33$ for 4Zo and $21 \leq N_{2} \leq 28$ for 4Zo-2MIo. However, the optimal number of $N_{1}$ analogues of the first level of 4Zo-2MIo varied to a greater extent: $48 \leq N_{1} \leq 84$. In this latter method, it may be problematic to consider a fixed and unique value for all regions.

As for the pressure levels, Z1000 that was previously systematically selected twice (Table \ref{table:levels_GA_4Zo}) was here less often chosen (once or even not at all) for both 4Zo-2MIo (Table \ref{table:levels_GA_4Zo_2MIo}) and 4Zo-4MIo (not shown). There was indeed a vertical shift of the previously selected Z1000 for higher levels, that was even slightly stronger with four moisture predictors than with two. This change is likely due to the fact that when considering only the circulation analogy, the method tried to take into account information that can serve as proxy for moisture assessment, whereas it did not need it with the moisture index. This aspect has never been demonstrated before, as sequential calibration tools do not allow it. It can only be assessed by a global optimization technique that can work jointly on both levels of analogy.

The selected pressure levels for the analogy on the moisture index were strongly centered around 700~hPa and 600~hPa. No other value was selected when considering two pressure levels (Table \ref{table:levels_GA_4Zo_2MIo}), and when considering four levels, 850~hPa and 500~hPa were sometimes also selected (not shown). However, even in this latter method, the 700~hPa and 600~hPa levels still held 78 \% of the selection. It was thus more efficient, in terms of prediction performance, to consider one of this level several times at different hours rather than selecting another pressure level. Besides, the optimizer never chose the same pressure level at the same hour for any variable, even though it was allowed to do so. The selected pressure levels for the analogy on the moisture index differed from the parameters resulting from the reference method optimized by the sequential procedure (Tables \ref{table:params_R2} and \ref{table:levels_GA_4Zo_2MIo}, last row). It is possible that the level 850~hPa is more optimal for the region for which the reference parameterization was established. However, this selection is usually not reconsidered when applying the sequential procedure.

The selection of the temporal windows for the atmospheric circulation was similar to the preceding optimization (in the order of increasing pressure: 12~h, 24/30~h, 12~h, 30~h), but sometimes with a bit more variability. When it comes to the moisture analogy, there was a clear trend to select 12~h and 24~h when considering two predictors. However, it must be remembered that this holds for our predictand, the accumulated precipitation over 06-30~h~UTC, and that it is expected to differ if we change the temporal window (e.g. 00-24~h~UTC) or its duration.

The optimized spatial windows for the atmospheric circulation have also changed (Figure \ref{fig:spatial_windows_4Zo-2MIo}; results of 4Zo-4MIo are not shown but present the same trends). The very large domains on the upper level of the four first subregions were not present anymore, and more variability could be observed. The selected points for the analogy on the moisture index were always located nearby the catchment, including at least one of the nearest points from the reanalysis dataset, and the spatial windows were relatively small. Thus, for this case study, there is no need to look for distant moisture information and the search could be reduced to a smaller domain.

The CRPSS scores were improved by considering the moisture information (Figure \ref{fig:figure_crpss_4Zo-2HIo} to be compared with Figure \ref{fig:figure_crpss_4Zo}). The optimized methods also performed significantly better than the reference method based on the moisture index optimized by the sequential procedure. However, there was no drastic difference between both 4Zo-2MIo and 4Zo-4MIo methods, which suggests that considering four moisture predictors is not necessary. When it comes to improvements for days with precipitation above the three thresholds (P\(\geq\)1~mm, P\(\geq\)0.1\(\cdot\)P10 and P\(\geq\)0.5\(\cdot\)P10), the conclusion is the same as before, that is a significant improvement of the prediction compared to the reference method, mainly for heavy rainfall.

The analysis of the parameters cross-compatibility (Figure \ref{fig:crossing_4Zo-2MIo}) was also very similar to the one on the circulation analogy only (results for 4Zo-4MIo not shown as very similar). The same pattern could be observed, with a drop of performance for the subregions submitted to different meteorological influences. However, the losses of performance were globally more important than before, suggesting that more complex methods with moisture variables are less transposable to another subregion (consistent with the observations of \citet{Chardon2014}), even though both were located within the same grid cell of the reanalysis dataset. Moisture fields have greater variability than pressure fields, and thus a change in the spatial windows can have a greater impact on the method performance. Indeed, the two regions with the lowest cross-compatibility with the others were the Upper Rhone Valley (7) and the Southeast ridges (8), which had similar optimal pressure levels and temporal windows to other regions, but had rather different spatial windows on the moisture predictor.

Predictors based on moisture variables do significantly increase the prediction skill, and are thus recommended, as long as they are reliable. In real-time forecasting, their reliability depends on the lead time: for lead times superior to 3-4 days the uncertainties related to moisture variables from NWP models becomes fairly high, which reduces the relevance of methods relying on this information. In climate downscaling studies, it mainly depends on the coherence of the climatologies between the archive and the GCM model outputs. One should however not establish an AM with moisture variables for a too large region, as the transferability is reduced \citep[see also][]{Chardon2014}.

@@

The optimization of the AM by means of GAs has been undertaken in successive stages by releasing progressively new degrees of freedom. This approach allowed us to differentiate the contributions to performance gains, as well as to identify possible over-parametrization. The main improvements obtained in the present case study are due to the following factors:

\begin{itemize}
	\item Using four pressure levels for the analogy of circulation seemed to be an optimal number for the studied region, length of archive available, and target predictand considered. Beyond four, the validation score was more variable, revealing a loss in robustness due to over-parametrization.
	\item The automatic and joint optimization of all parameters: the analogues number, the selection of the pressure levels and the temporal windows, and the spatial windows. These parameters are highly interdependent, so one needs to optimize them jointly in order to identify optimal combinations. Indeed, there is a strong interdependence between space and time in the atmospheric circulation, so that e.g. the spatial window should move upstream the main atmospheric flow for earlier temporal windows.	Traditional calibration procedures based on a systematic assessment of every combination is not possible anymore when considering more than two pressure levels.
	\item The introduction of distinct spatial windows between pressure levels. The synoptic circulation is characterized by features with very different scales depending on the height, and important information for predicting rainfall is not necessarily located in the same area from one level to another. However, the optimized spatial windows are consistent in between the subregions.
	\item The weighting of the analogy criteria between different pressure levels. It can be influenced by the variability of the geopotential with altitude, and the change of some levels significance with the targeted region. There was a trend for the weighting of circulation predictors to decrease with the increase in pressure, as one can see in Figure \ref{fig:levels_weights_average} for the three optimized methods. However, the values stayed around equity. This may not be the most influencing factor, and we may suggest to remove it first when trying to reduce the number of degrees of freedom.
	\item The joint optimization of the circulation and moisture analogy levels, that are usually calibrated successively. We have been able to demonstrate that there is a dependency between the analogy levels, and that in order to approach the optimal parameters, one must consider them jointly.
\end{itemize}

GAs have proved very useful to optimize complex variants of the AM, and to assess new degrees of freedom that were not available so far. However, it can be dangerous to add too many parameters to optimize. Indeed, the optimizer will probably use them to successfully improve the calibration score, but the validation control remains very important in order to determine if one is actually improving the method, or if it is being over-parametrized. Moreover, it might not always be desirable to increase the number of degrees of freedom, and some constraints (e.g. same weighting of the analogy criteria between different pressure levels) can be justified. However, one should first assess the consequence of a constraint before establishing it. In this sense, even though not all degrees of freedom can be found useful, GAs allow assessing their influence.

The convergence of parallel optimizations decreased when the method to optimize becomes more and more complex. The optimizer does not always converge to the exact global optimum, but to its surroundings. This is related to the fact that the optimization slows down when it gets closer to the global optimum, and that one has to stop it before the end, due to the required processing time (see Figure \ref{fig:evolution} as example). The resulting parameters might sometimes present non negligible differences, even though the score was almost similar. Through some Monte-Carlo analyses of the parameter space properties of the AM, \citet{Horton2012a} showed that some parameters of the method have a wide range of acceptable values. The spatial windows, for example, can be larger than the optimal size without much impact on the score, while they cannot be smaller \citep[see also][]{Bontron2004}. We could also observe that the selection of the pressure level is not a parameter as discrete as we would have thought, and that choosing another level may have reduced consequences on the performance. This is particularly true for higher pressure levels but can be more critical for lower layers. It was thus interesting to sometimes get several sets of near-optimal parameters, but with some nuances, in order to get an idea of the sensitivity of the parameters for a given region, and to compare the score on the VP. In this regard, a cross-validation technique may be advisable. However, as solutions identified at different regions of the parameters space might provide sufficiently good performance, an ensemble of these could be used, instead of a unique solution. These could account for the parameters uncertainty of the AM and increase the sample size contributing to the empirical distribution of precipitation values. An approach that can also be recommended is to first explore a wide range of the parameter space with some optimizations, and to narrow it according to the results for more targeted optimizations that are likely to go faster and to perform better.

@@

Furthermore, AMs optimized with GAs showed an improvement of the prediction for days with heavier precipitation, including extremes. Even though no new extreme value is added to the existing precipitation archive, the distribution of analogue precipitation values for a target situation can move towards the targeted extreme by sampling better candidate situations. Then, the subset of precipitation values collected on the analogue dates can be considered as a sample of the conditional distribution of precipitation associated with this situation. A truncated exponential or a gamma distribution model can be fitted and extrapolated to extreme values not contained in the sample or even in the whole precipitation archive \citep{Obled2002}. Another possible approach is by combining AMs with other methods \citep[e.g.][]{Chardon2014}.

@@

Strong dependencies between the parameters of the AM could be observed. Thus, the sequential calibration, which optimizes the parameters successively, may not lead to the optimal combination. Moreover, it contains several manual systematic assessment, such as the selection of the pressure levels and the temporal windows. GAs, however, can select the pressure levels and the time windows automatically, which can save a considerable amount of human time. A great advantage of a global optimization is its ability to approach or reach optimal parameter values when they are considered jointly.

A dependence in the selected parameters between the analogy of circulation and the moisture variables was identified. When the two levels are considered together, the optimal parameters of the circulation analogy were different. This complexity can only be exploited in a suitable manner by global optimization methods.

For the present case study, there seemed to be an optimum number of pressure levels to consider for the analogy of circulation, which is four, before losing consistency of the real gains. The analogy of circulation was improved by introducing a weighting between pressure levels, and considering independent spatial windows between pressure levels.

GAs provided parameterizations of AMs that exceeded the performance of the sequential calibration. In addition, it has been observed that the prediction for days with strong precipitation were improved to a greater extent, which is clearly interesting in the framework of flood forecasting.

This work is by far not exhaustive and means to open a door for new explorations of AMs with GAs. It is even possible to let the optimizer chose the meteorological variable to be used as predictor, as well as the analogy criteria, which is the topic of work in progress. Moreover, the AM has been explored for decades for precipitation forecasting, but not as intensively for other predictands. A global optimizer, such as GAs, can speed up this assessment significantly.

@@

\begin{figure}[t]
	\centerline{\includegraphics[width=8.8cm]{figures/fig02.pdf}}
	\caption{Performance score (CRPSS) on the CP and the VP for three stations (1 - Swiss Chablais, 5 - Southern valleys, 8 - Southeast ridges) when varying the number of geopotential height levels available to the optimizer.}
	\label{fig:figure_nb_levels}
\end{figure}

\begin{figure}[t]
	\centerline{\includegraphics[width=8.4cm]{figures/fig03.pdf}}
	\caption{Optimized spatial windows for the 4Zo method (analogy of atmospheric circulation on four pressure levels). The pressure levels are ordered by increasing pressure and increasing time for the same levels.}
	\label{fig:spatial_windows_4Zo}
\end{figure}

@@

\begin{figure}[t]
	\centerline{\includegraphics[width=8.4cm]{figures/fig06.pdf}}
	\caption{Optimized spatial windows for the 4Zo-2MIo method (analogy of atmospheric circulation on four pressure levels and the analogy on the moisture index on two pressure levels). The pressure levels are ordered by increasing pressure and increasing time for the same levels.}
	\label{fig:spatial_windows_4Zo-2MIo}
\end{figure}

@@

\begin{figure}[t]
	\centerline{\includegraphics[width=8.8cm]{figures/fig08.pdf}}
	\caption{Number of analogues for the different regions and the various methods resulting from the optimization. Method 4Zo is made of a single level of analogy with $N_{1}$ analogues, whereas 4Zo-2MIo has two levels of analogy with respectively $N_{1}$ and $N_{2}$ analogues.}
	\label{fig:figure_nb_analogs}
\end{figure}

@@

\begin{figure}[t]
	\centerline{\includegraphics[width=8.8cm]{figures/fig10.pdf}}
	\caption{Distribution of the optimal weights for the predictors of the first level of analogy (geopotential heights) of the 4Zo, 4Zo-2MIo, and 4Zo-4MIo methods. The results are those of the ten subregions. The geopotential heights are sorted by increasing pressure and hour.}
	\label{fig:levels_weights_average}
\end{figure}

@@

\begin{table}[t]
	\caption{Parameters of the reference method on the atmospheric circulation (2Z). The first column is the level of analogy (0 for preselection), then comes the meteorological variable and its hour of observation within the target day (temporal window). The criterion used for the current level of analogy is then provided, as well as the number of analogues.}
	\footnotesize
	\begin{center}
		\begin{tabular}{ccccc}
			\hline
			Level & Variable & Hour & Criteria & Nb \\ 
			\hline 
			0 & \multicolumn{4}{l}{$\pm 60$ days around the target date} \\
			\hline 
			\multirow{2}{*}{1} & Z1000 & 12~h & \multirow{2}{*}{S1} & \multirow{2}{*}{50} \\
			& Z500 & 24~h & & \\ 
			\hline 
		\end{tabular} 
	\end{center}
	\label{table:params_R1}
\end{table}

>>>>

Parallel analyses showed that the analogy of circulation is incomplete, and that geopotential heights still contain relevant information that can improve the statistical relationship. Therefore, a third, followed by a fourth circulation predictor were added (still only geopotential heights). There was no constraint on the predictors, so that the same pressure level could be selected more than once. Further improvements were found on the performance score, both for the CP and the VP, confirming that this additional information was beneficial for the quality of the prediction.

@@

The number of circulation predictors (still only geopotential heights) was then successively increased up to ten, considering the weighting of the analogy criteria values. The addition of circulation predictors globally improved the prediction skill (for both the CP and the VP) only up to four predictors (Figure \ref{fig:figure_nb_levels}). Afterwards, the score on the VP was more variable, showing eventually even a decrease, which revealed an over-parametrization of the method, and thus a lack of robustness. After four predictors, the score on the CP did not increase substantially, and could even present some local decrease due to increasing difficulty for the optimizer to converge. Selecting four circulation predictors (geopotential heights) was considered optimal for this case study, since the gain in CRPSS was significant, and the model remained relatively simple. It cannot be excluded that another number would prevail in another region than the upper Rh\^{o}ne catchment, under other meteorological conditions, or with another reanalysis dataset.

@@

The chosen method for the atmospheric circulation analogy, based on four circulation predictors (geopotential heights), and that will be named 4Zo (o for optimized), was made of the following degrees of freedom:

@@

Globally, the same cross-compatibility structure could be observed for the VP (Figure \ref{fig:crossing_4Zo} bottom), but in this case, minor improvements were occasionally observed when crossing the parameters, due to the presence of other events in the VP that might be better predicted by a different parameter set. The relatively small differences in score between parameterizations indicated that even though the parameters might differ significantly, the performance might not be drastically affected. Even a change in the pressure level did not mean a radical drop of the score value. A different parametrization may lead to a distinct selection of analogue days, and thus to an improvement of the prediction under certain weather conditions at the expense of others.

@@

Two methods were assessed: one with two moisture predictors (moisture index on two pressure levels or at two different hours) named 4Zo-2MIo, and one with four moisture predictors named 4Zo-4MIo. When introducing two predictors for the moisture analogy, the number of degrees of freedom raised to 42, and to 54 with four predictors. However, there was no substantial difference in the performance scores between both 4Zo-2MIo and 4Zo-4MIo methods, which suggests that considering four moisture predictors is not necessary. For this reason, only the results of 4Zo-2MIo are presented.

@@

\subsection{Results for the 4Zo-2MIo method}

@@

The CRPSS scores of the optimized 4Zo-2MIo method are provided in Figure \ref{fig:figure_crpss_4Zo-2HIo} and amounted in average to 40~\% (CP) and 40.3~\% (VP), compared to 35.2~\% (CP) and 36.2~\% (VP) for the reference method 2Z-2MI on the moisture analogy optimized with the sequential procedure. The parameters cross-compatibility are shown in Figure \ref{fig:crossing_4Zo-2MIo}. As for 4Zo, the 4Zo-2MIo method presented larger improvements in the prediction of significant rainfall (Table \ref{table:scores_thresholds_4Zo-2MIo}).

@@

When optimizing a method made of two levels of analogy, the introduction of moisture variables in the second level had an influence on the parameter values of the first level. It means that the two levels of analogy bring complementary information, and are thus not independent. This is first visible on the number $N_{1}$ of analogues to be selected on the first level, and on the selection of the pressure levels for the circulation analogy. If the change in the optimal value of $N_{1}$ was already known, a change of the optimal pressure levels for the circulation analogy has never been identified before.

As for the sequential procedure, the optimal value of $N_{1}$ increased when adding a second level of analogy (Figure \ref{fig:figure_nb_analogs}). One can also notice that the optimal number of analogues $N_{2}$ for the second level of analogy of 4Zo-2MIo was slightly inferior to $N_{1}$ from 4Zo, but very close. There is globally a common tendency between the optimal analogue number values of both methods: $N_{1}$ of the 4Zo method, and $N_{1}$ and $N_{2}$ of 4Zo-2MIo tend to be higher or lower altogether for a given region.

The optimal final numbers of analogues did not vary much: $23 \leq N_{1} \leq 33$ for 4Zo and $21 \leq N_{2} \leq 28$ for 4Zo-2MIo. However, the optimal number of the $N_{1}$ analogues of the first level of 4Zo-2MIo varied to a greater extent: $48 \leq N_{1} \leq 84$. In this latter method, it may be problematic to consider a fixed and unique value for all regions.

As for the pressure levels, Z1000 that was previously systematically selected twice (Table \ref{table:levels_GA_4Zo}) was here less often chosen (once or even not at all) for 4Zo-2MIo (Table \ref{table:levels_GA_4Zo_2MIo}). There was indeed a vertical shift of the previously selected Z1000 for higher levels, that was even slightly stronger with four moisture predictors than with two. This change is likely due to the fact that when considering only the circulation analogy, the method tried to take into account information that can serve as proxy for moisture assessment, whereas it did not need it with the moisture index. This can only be assessed by a global optimization technique that can work jointly on both levels of analogy.

The selected pressure levels for the analogy on the moisture index were strongly centered around 700~hPa and 600~hPa. No other value was selected when considering two moisture predictors (Table \ref{table:levels_GA_4Zo_2MIo}). It was sometimes more efficient, in terms of prediction performance, to consider the moisture at 700~hPa twice, but at different hours, rather than selecting another pressure level. Besides, the optimizer never chose the same pressure level at the same hour for any variable, even though it was allowed to do so. The selected pressure levels for the moisture analogy differed from the reference method (Tables \ref{table:params_R2} and \ref{table:levels_GA_4Zo_2MIo}, last row).

The selection of the temporal windows for the atmospheric circulation was similar to the preceding optimization (in the order of increasing pressure: 12~h, 24/30~h, 12~h, 30~h), but sometimes with some variability. When it comes to the moisture analogy, there was a clear tendency to select 12~h and 24~h. However, it must be remembered that this holds for our predictand, the accumulated precipitation over 06-30~h~UTC, and that it is expected to differ if the temporal window changes (e.g. 00-24~h~UTC, or another accumulation duration).

The optimized spatial windows for the atmospheric circulation have also changed (Figure \ref{fig:spatial_windows_4Zo-2MIo}). The very large domains on the upper level of the four first subregions were not present anymore, and more variability could be observed. The selected points for the moisture analogy were always located nearby the catchment, including at least one of the nearest points from the reanalysis dataset, and the spatial windows were relatively small. Thus, for this case study, there is no need to look for distant moisture information and the search could be reduced to a smaller domain.

The CRPSS scores were improved by considering the moisture information (Figure \ref{fig:figure_crpss_4Zo-2HIo} to be compared with Figure \ref{fig:figure_crpss_4Zo}). The optimized method also performed significantly better than the 2Z-2MI reference method optimized by the sequential procedure. When it comes to improvements for days with precipitation above the three thresholds (P\(\geq\)1~mm, P\(\geq\)0.1\(\cdot\)P10 and P\(\geq\)0.5\(\cdot\)P10), the conclusion is the same as before, that is a significant improvement of the prediction compared to the reference method, mainly for heavy rainfall.

The analysis of the parameters cross-compatibility (Figure \ref{fig:crossing_4Zo-2MIo}) was also very similar to the one on the circulation analogy only. The same pattern could be observed, with a drop of performance for the subregions submitted to different meteorological influences. However, the losses of performance were globally more important than before, suggesting that more complex methods with moisture variables are less transposable to another subregion (consistent with the observations of \citet{Chardon2014}), even though both were located within the same grid cell of the reanalysis dataset. Moisture fields have greater variability than pressure fields, and thus a change in the spatial windows can have a greater impact on the method performance. Indeed, the two regions with the lowest cross-compatibility with the others were the Upper Rhone Valley (7) and the Southeast ridges (8), which had similar optimal pressure levels and temporal windows to other regions, but had rather different spatial windows on the moisture predictor.

Predictors based on moisture variables do significantly increase the prediction skill, and are thus recommended, as long as they are reliable. In real-time forecasting, their reliability depends on the lead time: for lead times superior to 3-4 days the uncertainties related to moisture variables from NWP models becomes fairly high, which reduces the relevance of methods relying on this information. In climate downscaling studies, it mainly depends on the coherence of the climatologies between the archive and the GCM model outputs. One should however not establish an AM with moisture variables for a too large region, as the transferability is reduced \citep[see][for alternative approaches]{Chardon2014}.

@@

The optimization of the AM by means of GAs has been undertaken in successive stages by releasing progressively new degrees of freedom. This approach allowed us to differentiate the contributions to performance gains, as well as to identify possible over-parametrization. The main improvements obtained in the present case study are due to the following elementss:

\begin{itemize}
	\item Using four pressure levels for the circulation analogy seemed to be an optimal number for the studied region, length of archive available, and target predictand considered. Beyond four, the validation score was more variable, revealing a loss in robustness due to over-parametrization.
	\item The automatic and joint optimization of all parameters: the analogues number, the selection of the pressure levels and the temporal windows, and the spatial windows. These parameters are highly interdependent, so one needs to optimize them jointly in order to identify optimal combinations. Indeed, there is a strong interdependence between space and time in the atmospheric circulation, so that e.g. the spatial window should move upstream the main atmospheric flow for earlier temporal windows.
	\item The introduction of distinct spatial windows between pressure levels. The synoptic circulation is characterized by features with very different scales depending on the height, and important information for predicting precipitation is not necessarily located in the same area from one level to another.
	\item The weighting of the analogy criteria between different pressure levels. It can be influenced by the variability of the geopotential height with altitude, or the levels significance in regards to the meteorological processes specific to a region. There is a trend for the weighting of circulation predictors to decrease with the increase in pressure, as one can see in Figure \ref{fig:levels_weights_average} for the three optimized methods. However, the values stayed around equity. This may not be the most influencing factor, and we may suggest to remove it first when trying to reduce the number of degrees of freedom.
	\item The joint optimization of the circulation and moisture analogy levels, that are usually calibrated successively. We have been able to demonstrate that there is a dependency between the analogy levels, and that in order to approach the optimal parameters, one must consider them jointly.
\end{itemize}

GAs have proved very useful to optimize complex variants of the AM, and to assess new degrees of freedom that were not available so far. However, it can be dangerous to add too many parameters to optimize. Indeed, the optimizer will probably use them to successfully improve the calibration score, so the validation control remains very important in order to determine if one is actually improving the method, or if it is being over-parametrized. Moreover, it might not always be desirable to increase the number of degrees of freedom, and some constraints (e.g. same weighting of the analogy criteria between different pressure levels) can be justified. However, one should first assess the consequence of a constraint before establishing it. In this sense, even though not all degrees of freedom can be found useful, GAs allow assessing their influence. Finally, GAs could be used in order to identify e.g. the best couples of spatial and temporal windows, to later create a simpler regional method.

The convergence of parallel optimizations decreased when the method to optimize became more and more complex. The optimizer did not always converge to the exact global optimum, but to its surroundings. This is related to the fact that the optimization slows down when it gets closer to the global optimum, and that one has to stop it before the end, due to the required processing time (see Figure \ref{fig:evolution} as example). The resulting parameters might sometimes present non negligible differences, even though the score was almost similar. Through some Monte-Carlo analyses of the parameter space properties of the AM, \citet{Horton2012a} showed that some parameters of the method have a wide range of acceptable values. The spatial windows, for example, can be larger than the optimal size without much impact on the score, while they cannot be smaller \citep[see also][]{Bontron2004}. We could also observe that the selection of the pressure level is not a parameter as discrete as we would have thought, and that choosing another level may have reduced consequences on the performance. This is particularly true for higher pressure levels but can be more critical for lower layers. It was thus interesting to sometimes get several sets of near-optimal parameters, but with some nuances, in order to get an idea of the sensitivity of the parameters for a given region, and to compare the score on the VP. In this regard, a cross-validation technique may be advisable. However, as solutions identified at different regions of the parameters space might provide sufficiently good performance, an ensemble of these could be used, instead of a unique solution. These could account for the parameters uncertainty of the AM and increase the sample size contributing to the empirical distribution of precipitation values. An approach that can also be recommended is to first explore a wide range of the parameter space with some optimizations, and to narrow it according to the results for more targeted optimizations that are likely to go faster and to perform better.

@@

Furthermore, AMs optimized with GAs showed an improvement of the prediction for days with heavier precipitation, including extremes. Even though no new extreme value was added to the existing precipitation archive, the distribution of analogue precipitation values for a target situation can move towards the targeted extreme by sampling better candidate situations. Then, the subset of precipitation values collected on the analogue dates can be considered as a sample of the conditional distribution of precipitation associated with this situation. A truncated exponential or a gamma distribution model can be fitted and extrapolated to extreme values not contained in the sample or even in the whole precipitation archive \citep{Obled2002}. Another possible approach is by combining AMs with other methods \citep[e.g.][]{Chardon2014}.

@@

Strong dependencies between the parameters of the AM could be observed. Thus, the sequential calibration, which optimizes the parameters successively, may not lead to the optimal combination. Moreover, it contains several manual systematic assessment, such as the selection of the pressure levels and the temporal windows. GAs, however, can select the pressure levels and the temporal windows automatically, which can save a considerable amount of human time. A great advantage of a global optimization is its ability to approach or reach optimal parameter values when they are considered jointly.

A dependence in the selected parameters between the circulation analogy and the moisture variables was identified. When the two analogy levels are considered together, the optimal parameters of the circulation analogy changed. This complexity can only be exploited in a suitable manner by global optimization methods.

For the present case study, there seemed to be an optimum number of pressure levels to consider for the circulation analogy, which is four, before losing consistency of the real gains. The circulation analogy was improved by introducing a weighting between pressure levels, and considering independent spatial windows between pressure levels.

GAs provided parameterizations of AMs that exceeded the performance of the sequential calibration. In addition, it has been observed that the prediction for days with strong precipitation were improved to a greater extent, which is clearly interesting in the context of flood forecasting.

This work is by far not exhaustive and means to open a door for new explorations of AMs with GAs. It is even possible to let the optimizer chose the meteorological variable to be used as predictor, as well as the analogy criteria, which is the topic of work in progress. Moreover, the AM has been explored for decades for precipitation prediction, but not as intensively for other predictands. A global optimizer, such as GAs, can speed up this assessment significantly.

@@

\begin{figure}[t]
	\centerline{\includegraphics[width=8.8cm]{figures/fig02.pdf}}
	\caption{Performance score (CRPSS) on the CP and the VP for three stations (1 - Swiss Chablais; 5 - Southern valleys; 8 - Southeast ridges) when varying the number of geopotential height levels available to the optimizer.}
	\label{fig:figure_nb_levels}
\end{figure}

\begin{figure}[t]
	\centerline{\includegraphics[width=8.4cm]{figures/fig03.pdf}}
	\caption{Optimized spatial windows for the 4Zo method (analogy of the atmospheric circulation on four pressure levels).}
	\label{fig:spatial_windows_4Zo}
\end{figure}

@@

\begin{figure}[t]
	\centerline{\includegraphics[width=8.4cm]{figures/fig06.pdf}}
	\caption{Optimized spatial windows for the 4Zo-2MIo method (analogy of atmospheric circulation on four pressure levels and the analogy on the moisture index on two pressure levels).}
	\label{fig:spatial_windows_4Zo-2MIo}
\end{figure}

@@

\begin{figure}[t]
	\centerline{\includegraphics[width=8.8cm]{figures/fig08.pdf}}
	\caption{Optimal numbers of analogues for the different regions and the two methods, resulting from the optimization. Method 4Zo is made of a single level of analogy with $N_{1}$ analogues, whereas 4Zo-2MIo has two levels of analogy with respectively $N_{1}$ and $N_{2}$ analogues.}
	\label{fig:figure_nb_analogs}
\end{figure}

@@

\begin{figure}[t]
	\centerline{\includegraphics[width=8.8cm]{figures/fig10.pdf}}
	\caption{Distribution of the optimal weights for the predictors of the first level of analogy (geopotential heights) of the (blue) 4Zo, (red) 4Zo-2MIo, and (green) 4Zo-4MIo methods. The results are those of the ten subregions. The geopotential heights are sorted by increasing pressure and hour.}
	\label{fig:levels_weights_average}
\end{figure}

@@

\begin{table}[t]
	\caption{Parameters of the reference method on the atmospheric circulation (2Z). The first column is the level of analogy (0 for preselection), then comes the meteorological variable and its hour of observation within the target day (temporal window). The criterion used for the current level of analogy is then provided, as well as the number of analogues.}
	\footnotesize
	\begin{center}
		\begin{tabular}{ccccc}
			\hline
			Level & Variable & Hour & Criterion & Nb \\ 
			\hline 
			0 & \multicolumn{4}{l}{$\pm 60$ days around the target date} \\
			\hline 
			\multirow{2}{*}{1} & Z1000 & 12~h & \multirow{2}{*}{S1} & \multirow{2}{*}{50} \\
			& Z500 & 24~h & & \\ 
			\hline 
		\end{tabular} 
	\end{center}
	\label{table:params_R1}
\end{table}

\begin{table}[t]
	\caption{Parameters of the reference method with moisture variables (2Z-2MI). Same conventions as Table \ref{table:params_R1}}
	\footnotesize
	\begin{center}
		\begin{tabular}{ccccc}
			\hline 
			Level & Variable & Hour & Criterion & Nb \\ 
			\hline 
			0 & \multicolumn{4}{l}{$\pm 60$ days around the target date} \\
			\hline 
			\multirow{2}{*}{1} & Z1000 & 12~h & \multirow{2}{*}{S1} & \multirow{2}{*}{70} \\
			& Z500 & 24~h & & \\ 
			\hline
			\multirow{2}{*}{2} & TPW * RH850 & 12~h & \multirow{2}{*}{RMSE} & \multirow{2}{*}{30} \\
			& TPW * RH850 & 24~h & & \\ 
			\hline 
		\end{tabular} 
	\end{center}
	\label{table:params_R2}
\end{table}
